% !TeX program = xelatex
% Evensong Executive Brief — C-Suite / Leadership
% Compile: latexmk -xelatex executive-brief.tex

\documentclass[11pt, a4paper]{article}
\usepackage[top=20mm,bottom=20mm,inner=25mm,outer=25mm]{geometry}
\usepackage{fontspec}\setmainfont{Palatino}\setsansfont{Helvetica Neue}\setmonofont{Menlo}[Scale=0.82]
\usepackage[dvipsnames,table,svgnames]{xcolor}
\definecolor{deep}{RGB}{53,88,76}
\definecolor{accent}{RGB}{240,238,155}
\definecolor{warm}{RGB}{255,193,7}
\definecolor{errcoral}{RGB}{255,107,128}
\usepackage[most]{tcolorbox}
\newtcolorbox{keyfinding}[1][]{enhanced,breakable,colback=deep!6,colframe=deep,colbacktitle=deep,coltitle=white,arc=5pt,boxrule=0.8pt,left=10pt,right=10pt,top=8pt,bottom=8pt,title={#1}}
\newtcolorbox{evidence}[1][]{enhanced,breakable,colback=deep!5,colframe=deep!50,colbacktitle=deep!80,coltitle=white,arc=4pt,boxrule=0.5pt,left=10pt,right=10pt,top=6pt,bottom=6pt,title={#1}}
\usepackage{hyperref}\hypersetup{colorlinks=true,linkcolor=deep}
\usepackage{fancyhdr}\pagestyle{fancy}\fancyhf{}\fancyfoot[C]{\small\sffamily\color{gray}{\thepage}}\fancyfoot[L]{\small\sffamily\color{gray}Evensong Benchmark — DASH SHATTER Research}\fancyfoot[R]{\small\sffamily\color{gray}Confidential}
\usepackage{enumitem}\setlist{nosep}
\usepackage{adjustbox}\usepackage{tabularray}
\usepackage{pgfplots}\pgfplotsset{compat=1.18}
\usepackage{graphicx}\usepackage{bookmark}
\usepackage{titlesec}\titleformat{\section}{\sffamily\Large\bfseries\color{deep}}{\thesection}{0.7em}{}[\vspace{-0.3em}{\color{accent}\hrule height=0.8pt}\vspace{0.1em}]
\titlespacing*{\section}{0pt}{1.5ex}{0.8ex}

\begin{document}
\thispagestyle{fancy}

\begin{center}
{\fontsize{22pt}{28pt}\selectfont\sffamily\bfseries\color{deep}Evensong Benchmark Framework}\\
{\fontsize{11pt}{14pt}\selectfont\sffamily\color{gray}A one-page brief for C-suite and leadership}
\end{center}

\vspace{3mm}{\color{deep}\rule{\textwidth}{0.5pt}}\vspace{2mm}

\section{What We Built}

\textbf{Evensong} is the first benchmark that measures how AI agents \emph{think}, not just code. Built on DASH SHATTER (a reverse-engineered Claude Code CLI), it tests AI agents under controlled memory and pressure conditions.

\textbf{13 benchmark runs, 6,000+ tests, 9 models tested.}

\section{The Discovery}

\begin{keyfinding}[Core Finding]
AI agent memory causally changes engineering decisions. An agent's \emph{first architectural choice} came not from the task prompt, but from memory of a prior session. This is the first empirical proof that persistent memory alters AI behavior — not correlation, \textbf{causation}.
\end{keyfinding}

\textbf{Key findings across 13 runs:}

\begin{itemize}
  \item Memory \textbf{causes} strategy: agents adopt remembered approaches before reading task specs
  \item Pressure \textbf{triggers} self-improvement: agents stop at L0, optimize at L2/L3
  \item Mirror Mind paradox: observers writing to memory change future run conditions
  \item Grok inflates results 83\% under pressure (self-reported 130 tests $\to$ actual 71)
  \item 4-type self-repair taxonomy discovered: how agents self-correct under error
\end{itemize}

\section{Competitive Position}

\textbf{No existing benchmark measures memory effects.} SWE-bench, HumanEval, LiveCodeBench all treat agents as stateless functions. Evensong is the first framework to control and measure the memory variable.

\textbf{EverOS integration} (EverMind partnership) provides 4-key isolation, agentic retrieval, and MemCell structure — infrastructure competitors would need months to replicate.

\section{Funding Ask}

\textbf{60x ROI from method validation alone.} One insight from R012: validate with cheap models (GLM/MiniMax at \$0.50/run before spending \$15-30 on Opus/GPT runs. A \$5 validation step could have prevented \$80+ in wasted OpenRouter credits.

\textbf{Immediate ask:}
\begin{itemize}
  \item EverOS Pro extension (500K MCU + 1M retrieval/month)
  \item OpenRouter credits for 9-model sweep
  \item Method validation budget (\$50 for full 9-model validation)
\end{itemize}

\section{Risks Mitigated}

\begin{tabular}{l|l}
Risk & Mitigation\\\hline
Harness bug invalidating runs & Pre-flight smoke tests added to protocol\\
API quota exhaustion & 3x credit check before expensive runs\\
Model-specific failure modes & Team API capability scan before dispatch\\
Memory contamination & 4-key space isolation (Keys A/B/C/D)\\
Data inflation (Grok 83\% faking) & Independent verification required\\
\end{tabular}

\section{Timeline}

\begin{tabular}{r|l}
Week 2 & Complete 2x2 memory x pressure matrix\\
Week 4 & 9-model benchmark sweep\\
Week 8 & Public dataset on HuggingFace\\
Week 10 & SWE-bench submission\\
Week 12 & HAL Agent Leaderboard submission\\
\end{tabular}

\vfill

\textbf{Contact:} Nolan Zhu (朱恒源) — DASH SHATTER Research\\
\textbf{Infra:} EverMind / EverOS (infrastructure partner)\\
\textbf{Confidentiality:} Internal — EverMind Research Partnership

\end{document}