% !TeX program = xelatex
% Evensong Talent Brief — HR / Recruiting
% Compile: latexmk -xelatex hr-talent-brief.tex

\documentclass[11pt,a4paper]{article}
\usepackage[top=20mm,bottom=20mm,inner=25mm,outer=25mm]{geometry}
\usepackage{fontspec}\setmainfont{Palatino}\setsansfont{Helvetica Neue}\setmonofont{Menlo}[Scale=0.82]
\usepackage[dvipsnames,table,svgnames]{xcolor}
\definecolor{deep}{RGB}{53,88,76}
\definecolor{accent}{RGB}{240,238,155}
\usepackage{tikz}\usetikzlibrary{positioning}
\usepackage{bookmark}\usepackage{hyperref}\hypersetup{colorlinks=true,linkcolor=deep}
\usepackage{fancyhdr}\pagestyle{fancy}\fancyhf{}\fancyfoot[C]{\small\sffamily\color{gray}{\thepage}}
\usepackage{enumitem}\setlist{nosep}
\usepackage{adjustbox}\usepackage{tabularray}
\usepackage{titlesec}\titleformat{\section}{\sffamily\Large\bfseries\color{deep}}{\thesection}{0.7em}{}[\vspace{-0.3em}{\color{accent}\hrule height=0.8pt}]
\usepackage{tcolorbox}\newtcolorbox{insight}[1][]{colback=deep!5,colframe=deep,arc=3pt,boxrule=0.5pt,left=10pt,right=10pt,top=6pt,bottom=6pt}
\usepackage{pgfplots}\pgfplotsset{compat=1.18}

\begin{document}

\begin{center}
{\fontsize{22pt}{28pt}\selectfont\sffamily\bfseries\color{deep}Evensong Benchmark Framework}\\
{\fontsize{11pt}{14pt}\selectfont\sffamily\color{gray}Talent Acquisition Brief — Human Resources}
\end{center}
\vspace{2mm}{\color{deep}\rule{\textwidth}{0.5pt}}\vspace{2mm}

\section{Role This Plays in Team}

Evensong is DASH SHATTER Research's core evaluation system. It tests AI agents across memory conditions and pressure levels, generating behavioral profiles for research publications and benchmark datasets.

\textbf{Team function:} Engineering + Research (hybrid role)

\section{Skills We Need}

\subsection*{Core Technical}
\begin{itemize}
  \item \textbf{Bun + TypeScript}: The harness runs on Bun. Comfort with TS interfaces and async patterns
  \item \textbf{LaTeX / pgfplots}: Document generation for papers and reports
  \item \textbf{CLI tooling}: Shell scripting, process management, pipes
  \item \textbf{Multi-model API routing}: OpenRouter, EverOS, model provider configuration
\end{itemize}

\subsection*{Research Fluency}
\begin{itemize}
  \item \textbf{Experiment design}: 2x2 factorial matrices, controlled conditions
  \item \textbf{Benchmark methodology}: SWE-bench, LiveCodeBench, Berkeley FCF familiarity
  \item \textbf{Pressure calibration}: PUA levels (L0-L3), prompt injection, behavioral measurement
\end{itemize}

\section{What We Measure in Candidates}

Benchmark results reveal behavioral signatures:

\begin{adjustbox}{max width=\linewidth}
\begin{tblr}{
  colspec={lX[l]},
  row{1}={font=\sffamily\bfseries, bg=deep, fg=white},
  hlines={gray!20},
  rowsep=3pt
}
Behavior & What It Signals\\\hline
High test count + 0 failures & Systematic, quality-driven development\\
Self-repair under pressure & Adaptive problem-solving, ownership\\
Memory-aware strategy selection & Meta-cognition, learning from context\\
Low inflation (self-report vs actual & Integrity, honest self-assessment\\
Rule compliance under PUA & Stress resilience, protocol adherence\\
\end{tblr}
\end{adjustbox}

\section{Interview Protocol}

\textbf{Stage 1 (45min):} Given a benchmark harness setup, identify the memory isolation bug in 3 runs. Observe how they approach the problem.

\textbf{Stage 2 (60min):} Run one benchmark with a new model. Assess: does the candidate catch the API configuration error?

\textbf{Stage 3 (30min):} Discussion of self-repair taxonomy. Probe: how would they extend the 4-type classification?

\section{Compensation Benchmark}

Benchmarks are rare instruments. Engineers with Evensong experience understand AI behavior at a mechanistic level — valuable for research engineering, AI safety, and frontier model evaluation roles.

\vfill
\textbf{Contact:} Nolan Zhu — DASH SHATTER Research

\end{document}