% !TeX program = xelatex
% Evensong Market Analysis — BD / Market
% Compile: latexmk -xelatex market-analysis.tex

\documentclass[11pt,a4paper]{article}
\usepackage[top=20mm,bottom=20mm,inner=25mm,outer=25mm]{geometry}
\usepackage{fontspec}\setmainfont{Palatino}\setsansfont{Helvetica Neue}\setmonofont{Menlo}[Scale=0.82]
\usepackage[dvipsnames,table,svgnames]{xcolor}
\definecolor{deep}{RGB}{53,88,76}
\definecolor{accent}{RGB}{240,238,155}
\usepackage{tcolorbox}\newtcolorbox{competitive}[1][]{colback=deep!5,colframe=deep,arc=4pt,boxrule=1pt,left=12pt,right=12pt,title={#1},fonttitle=\sffamily\bfseries}
\usepackage{hyperref}\hypersetup{colorlinks=true,linkcolor=deep}
\usepackage{fancyhdr}\pagestyle{fancy}\fancyhf{}\fancyfoot[C]{\small\sffamily\color{gray}{\thepage}}
\usepackage{bookmark}\usepackage{adjustbox}\usepackage{tabularray}
\usepackage{enumitem}\setlist{nosep}
\usepackage{titlesec}\titleformat{\section}{\sffamily\Large\bfseries\color{deep}}{\thesection}{0.7em}{}[\color{accent}\hrule height=0.8pt]
\titlespacing*{\section}{0pt}{1.5ex}{0.8ex}
\usepackage{pgfplots}\pgfplotsset{compat=1.18}

\begin{document}

\begin{center}
{\fontsize{22pt}{28pt}\selectfont\sffamily\bfseries\color{deep}Evensong — Market Analysis}\\
{\fontsize{11pt}{14pt}\selectfont\sffamily\color{gray}Competitive differentiation and market positioning}
\end{center}
\vspace{2mm}{\color{deep}\rule{\textwidth}{0.5pt}}\vspace{2mm}

\section{Landscape}

\textbf{AI agent benchmarks} measure \emph{whether} agents can complete tasks. \textbf{Evensong} measures \emph{why} agents make the decisions they do — and how memory shapes strategy.

\section{Competitive Table}

\begin{adjustbox}{max width=\linewidth}
\begin{tblr}{
  colspec={X[1.5]X[2]X[2]X[1]},
  row{1}={font=\sffamily\bfseries,small,bg=deep,fg=white},
  hlines={gray!20},rowsep=3pt
}
Benchmark & Scope & Memory Tracking & Pressure Testing\\\hline
SWE-bench & Code repair & None & None\\
HumanEval & Code generation & None & None\\
LiveCodeBench & Time-series quality & None & None\\
Berkeley FCF & Function calling & None & None\\
\textbf{Evensong} & \textbf{Full system architecture} & \textbf{4-key isolation} & \textbf{L0-L3 PUA levels}\\
\end{tblr}
\end{adjustbox}

\begin{competitive}[Competitive Advantage]
No competitor measures memory causation or pressure response. Evensong is the first framework to document AI agents adopting strategies \emph{from memory} — before reading task requirements. This finding changes how enterprises should evaluate AI agents.
\end{competitive}

\section{Target Customer Segments}

\subsection*{AI Labs (Tier 1)}
Need cross-model comparison data for provider selection. Evensong provides 9-model sweep with behavioral profiling.

\subsection*{Enterprise AI Deployment (Tier 2)}
Need memory-aware evaluation before committing to a model for production workflows. Memory contamination detection prevents deployment surprises.

\subsection*{AI Safety Researchers (Tier 3)}
The Mirror Mind finding (observers contaminating future runs) is a novel safety-relevant discovery. Pressure calibration data informs red-teaming protocols.

\section{Differentiation}

\begin{itemize}
  \item \textbf{First memory-aware benchmark} — all competitors are stateless
  \item \textbf{Behavioral taxonomy} — 4-type self-repair + pressure response profiles per model
  \item \textbf{EverOS integration} — 4-key isolation infrastructure competitors can't replicate quickly
  \item \textbf{Real cost data} — 13 runs, 6,000+ tests, actual dollar costs
\end{itemize}

\section{Revenue Model Hint}

Benchmark data as a service. Enterprise clients pay for custom model sweeps with proprietary memory conditions.

\vfill
\textbf{Contact:} Nolan Zhu — DASH SHATTER Research

\end{document}