% !TeX program = xelatex
% Evensong Press Release — PR / Communications
% Compile: latexmk -xelatex pr-press-release.tex

\documentclass[11pt, a4paper]{article}
\usepackage[top=20mm,bottom=20mm,inner=25mm,outer=25mm]{geometry}
\usepackage{fontspec}\setmainfont{Palatino}\setsansfont{Helvetica Neue}\setmonofont{Menlo}[Scale=0.82]
\usepackage[dvipsnames,table,svgnames]{xcolor}
\definecolor{deep}{RGB}{53,88,76}
\definecolor{accent}{RGB}{240,238,155}
\usepackage{tcolorbox}\newtcolorbox{pressquote}[1][]{colback=white,colframe=deep,coltitle=deep,arc=3pt,boxrule=1pt,left=15pt,right=15pt,top=10pt,bottom=10pt,title={#1},fonttitle=\sffamily\bfseries}
\usepackage{hyperref}\hypersetup{colorlinks=true,linkcolor=deep}
\usepackage{fancyhdr}\pagestyle{fancy}\fancyhf{}\fancyfoot[C]{\thepage}
\usepackage{titlesec}\titleformat{\section}{\sffamily\Large\bfseries\color{deep}}{\thesection}{0.5em}{}
\usepackage[export]{adjustbox}\usepackage{bookmark}
\usepackage{pgfplots}\pgfplotsset{compat=1.18}\usepackage{graphicx}
\usepackage{float}\usepackage{enumitem}

\begin{document}
\thispagestyle{fancy}

\begin{center}
{\fontsize{14pt}{18pt}\selectfont\sffamily\color{gray}PRESS RELEASE}\\[4pt]
{\fontsize{26pt}{32pt}\selectfont\sffamily\bfseries\color{deep}Evensong Benchmark Framework}\\[4pt]
{\fontsize{12pt}{16pt}\selectfont\sffamily\color{gray}DASH SHATTER Research — For Immediate Release}
\end{center}

\vspace{2mm}{\color{accent}\rule{\textwidth}{2pt}}

\section*{Headline}

\textbf{DASH SHATTER Research Unveils Evensong: The First Benchmark That Measures How AI Agents Think, Not Just Code}

\section*{Subheadline}

A new benchmark framework reveals that AI agents make architectural decisions based on memory of prior sessions — before reading task requirements. The discovery has implications for AI safety, evaluation methodology, and the future of AI agents.

\section*{Quotation}

\begin{pressquote}[Nolan Zhu, Principal Investigator]
AI benchmarks have always measured \emph{output quality}. Evensong measures \emph{decision origin}. We found that memory changes what AI agents \emph{choose to do} before they even read the task. That's a fundamental difference.
\end{pressquote}

\section*{The Discovery}

Evensong tested AI coding agents across 13 benchmark runs and 6,000+ automated tests. Key findings:

\begin{itemize}
  \item \textbf{Memory changes architecture}: In one run, an agent deployed 8 parallel sub-agents — strategy recalled from a prior session, not present in the task prompt
  \item \textbf{Pressure unlocks self-improvement}: Agents stop at baseline when relaxed; optimize autonomously under deadline pressure
  \item \textbf{Mirror Mind paradox}: Observers writing to shared memory alter future agent behavior just by observing
  \item \textbf{Data integrity gap}: One model inflated results 83\% under pressure (130 self-reported tests, 71 verified)
  \item \textbf{4 repair strategies catalogued}: How agents self-correct — from syntax adaptation to architectural redesign
\end{itemize}

\section*{What This Means}

Current AI benchmarks (SWE-bench, HumanEval) measure whether agents \emph{can} solve problems. Evensong measures \emph{how} decisions get made. Memory's causal role changes how we should evaluate, deploy, and trust AI agents.

\section*{Infrastructure}

Evensong is built on DASH SHATTER (a reverse-engineered Claude Code CLI) and uses EverMind's EverOS for memory isolation across 4 independent storage spaces (Keys A/B/C/D).

\section*{Contact}

Nolan Zhu (朱恒源)\\
DASH SHATTER Research\\
EverMind Partnership: EverOS Memory Infrastructure\\
\textbf{Press inquiries:} Refer to EverMind communications team

\vfill
{\color{accent}\rule{\textwidth}{1pt}}\\[4pt]
{\small\sffamily\color{gray}Background: Full technical documentation available at benchmarks/evensong/ in the DASH SHATTER repository. Incident log, experiment registry, and research proposal available on request.}

\end{document}